---
title: "individual2"
author: "Nicholas Heger"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
str(choco)
```
##### Searching for the Best Chocolates  
1. How are the ratings distributed?
```{r}
hist(choco$Rating)
boxplot(choco$Rating)
```
  
The ratings are skewed left, with most ratings above the halfway mark of 2.5. We can see from the boxplot that the median is around 3.3 and there are roughly 4 outliers all around or below a rating of 2.  

2. How does the number of ratings depend on year?
```{r}
barplot(table(choco$Date))
```
  
The number of ratings per year is generally increasing until 2015, and it generally decreases after that, with a notably few ratings in 2017.  

3. Ratings vs Cocoa Percent
```{r}
cor(choco$Cocoa_Percent, choco$Rating, method = c("pearson", "kendall", "spearman"))
plot(choco$Cocoa_Percent, choco$Rating, main='Rating vs Cocoa Percent')
```
The correlation coefficient between the percent cocoa and rating is very low indicating that there is not much correlation between the two variables. Looking at a graph of these two variables it is clear that there is very little correlation and no trend to the graph.  

4. Favorable and non-favorable characteristics
```{r}
ggplot(choco, aes(x = Characteristics1, y = Rating)) +
  geom_col() + coord_flip()
ordered <- choco[order(choco$Rating),]
reversed <- choco[order(-choco$Rating),]
extremes <- rbind(ordered[1:10,], reversed[1:10,])
ggplot(extremes, aes(x = Characteristics1, y = Rating)) +
  geom_bar(stat='summary') + coord_flip()
print(unique(extremes[1:10,'Characteristics1']))
print(unique(extremes[11:20,'Characteristics1']))
ratingByCharacteristic1 <- choco %>%
  group_by(Characteristics1) %>%
  summarise_at(vars(Rating), funs(mean(., na.rm=TRUE))) %>% 
  arrange(Rating)
print(head(ratingByCharacteristic1['Characteristics1']))
print(tail(ratingByCharacteristic1['Characteristics1']))
```
Above shows first a graph of the characteristics vs rating, however there are a ton of characteristics, so I made a smaller graph with just the top 10 and bottom 10 rated chocolates and their corresponding charactersitics. I also listed out the characteristics in the bottom 10, followed by those in the top 10 according to the graph. Next I averaged the ratings of each characteristic and put that in a new dataframe. I've printed the bottom and top few characteristics according to the averages, respectively.
