---
title: "Lab2 Individual Report"
author: "Isaac Tegelerr"
date: "9/8/2020"
output: html_document
---

```{r}
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
str(choco)
```

1. How are the ratings distributed? Create both a histogram and a boxplot and describe them. Don’t forget to mention outliers, if there are any.

```{r}
hist(choco$Rating, breaks=seq(1.00, 4.00, .25))
boxplot(choco$Rating)
```

There does not seem to be any outliers for the distribution

2. How does the number of ratings depend on the year? Draw a bar chart of the number of reports.

To do this we will add a column of 1s to the front of the data frame and then aggregate the data by the year, summing up the 1s

```{r}
temp <- cbind(count = 1, choco)
sums <- aggregate(x = temp$count, by=list(temp$Date), FUN=sum)
barplot(sums$x, names.arg = sums$Group.1, main="Number of Ratings Per Year" )
```
3. Do ratings depend on the cacao percentage of a chocolate bar? Create both numerical and graphical summaries, and explain.

```{r}
means <- setNames(aggregate(x = choco$Cocoa_Percent, by = list(choco$Rating), FUN=mean), c("Rating", "Cocoa_Percentage"))
means
maxs <-  setNames(aggregate(x = choco$Cocoa_Percent, by = list(choco$Rating), FUN=max), c("Rating", "Cocoa_Percentage"))
maxs
mins <-  setNames(aggregate(x = choco$Cocoa_Percent, by = list(choco$Rating), FUN=min), c("Rating", "Cocoa_Percentage"))
mins


plot(maxs, bg=24, ylim=c(40,100), xlab="Rating", ylab="Cocoa Percentage")
points(means, col="blue", pch=22, bg=24)
points(mins,  col="red", bg=24)
legend(1, 60, legend = c("Max", "Mean", "Min"), col=c("black", "blue", "red"), pch=c(1, 22, 1))
```
Generally speaking you can see that the lower rated chocolate bars have higher Cocoa Percentage